Corpus Analysis

### 2020-08-28 inst/analysis/corpus-analysis.Rmd
Updated corpus

```{r load corpus}
corpus <- 
  read_fst(CORPUS_DETAILS_FILE) %>%
  as_tibble() %>%
  mutate_at(
    vars(
      starts_with("runnable_code"),
      starts_with("runnable_files"),
      starts_with("package_"),
      starts_with("funs_"),
      "evals"
    ), 
    replace_na, 
    0
  )
```

```{r load kaggle}
kaggle_kernels <- as_tibble(read_fst(KAGGLE_KERNELS_FILE))
kaggle_log <- as_tibble(read_fst(KAGGLE_LOG_FILE)) %>%
  transmute(
    id=str_replace(Command, ".*/kaggle-run/[^/]+/([^/]+)/.*", "\\1"),
    exitval=Exitval,
    runtime=as.duration(JobRuntime)
  )

kaggle <- left_join(
  kaggle_kernels,
  kaggle_log,
  by="id"
)
```

```{r load runnable code metadata}
runnable_code_metadata <- read_fst(PACKAGE_RUNNABLE_CODE_FILE) %>%
  as_tibble() %>%
  mutate(package=basename(package)) %>%
  select(package, file=filename, code)

# TODO semi_join with traced-programs.fst
```

```{r load run metadata}
runs <- read_fst(PACKAGE_RUNS_FILE) %>%
  as_tibble() %>%
  mutate(
    package=basename(package),
    file=as.character(path("./", basename(dirname(file)), basename(file)))
  ) %>%
  semi_join(corpus, by="package")
```

@@

```{r compute corpus programs}
corpus_programs <-
  left_join(
    filter(runs, exitval==0) %>% select(package, file, time), 
    runnable_code_metadata %>% select(package, file, code),
    by=c("package", "file")
  ) %>%
  transmute(
    package,
    file,
    type=basename(dirname(file)),
    code, 
    runtime=time
  )
```

There will be some programs that were not strictly part of examples, tests or vignettes, or they do not have
counts for code size or for execution time.

```{r compute program inconsistencies}
corpus_programs_inconsistent <-
  corpus_programs %>%
  filter(
    !(type %in% c("examples", "tests", "vignettes")) |
    is.na(code) | 
    is.na(runtime) |
    code == 0
  )

corpus_programs_inconsistent %>% 
    my_datatable()
```

Since, there are just a few of them, we ignore them.

```{r filter out inconsistent programs}
corpus_programs <- anti_join(corpus_programs, corpus_programs_inconsistent, by=c("package", "file"))
```

```{r summary}
runnable_code <- 
  corpus_programs %>%
  mutate(files=1) %>%
  pivot_wider(names_from=type, values_from=c(files, code)) %>%
  select(-file) %>%
  group_by(package) %>%
  summarise_all(sum, na.rm=TRUE) %>%
  group_by(package) %>%
  mutate(
    files=files_examples + files_tests + files_vignettes,
    code=code_examples + code_tests + code_vignettes
  )

rs <- function(name, value) {
  stopifnot(is.numeric(value))
  bind_rows(
    r(name, sum(value)),
    r(paste(name, "mean"), mean(value)),
    r(paste(name, "median"), median(value)),
    r(paste(name, "min"), min(value)),
    r(paste(name, "max"), max(value))
  )
}

@@

  rs("package programs",         runnable_code$files),
  rs("package programs code",    runnable_code$code),
  rs("package programs runtime", as.duration(runnable_code$runtime)),

  rs("examples programs",  runnable_code$files_examples),
  rs("tests programs",     runnable_code$files_tests),
  rs("vignettes programs", runnable_code$files_vignettes),

  rs("examples code",  corpus_program_examples$code),
  rs("tests code",     corpus_programs_tests$code),
  rs("vignettes code", corpus_programs_vignettes$code),

  rs("examples runtime",  corpus_program_examples$runtime),
  rs("tests runtime",     corpus_programs_tests$runtime),
  rs("vignettes runtime", corpus_programs_vignettes$runtime),

@@

corpus %>%
  mutate(
    outlier=evals > PACKAGE_EVALS_OUTLIER,# | package_r_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(outlier, package, NA),
    code_size=cut(package_r_code, size_bins, labels=size_bins_lasb, include.lowest=TRUE)
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
#      y=funs_with_eval,
      y=evals,
      label=label,
      size=code_size
    )
  ) +
  geom_point(alpha=.4) +
  geom_text(size=3, check_overlap = T, vjust=-1.5, na.rm = TRUE) + 
  geom_hline(aes(yintercept=funs_with_eval_mean), linetype="dashed", color="red", size=0.2) +
  geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="red", size=0.2) +
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(trans=scales::pseudo_log_trans(), breaks=c(0, 1, 2, 5, 10, 20, 50, 100, 200, 500)) +
  scale_size_discrete(range = c(1, 6)) +
  theme(
    legend.position=c(0.2, 0.85),
    legend.box="horizontal",
    legend.box.background = element_rect(fill="white", size=0.1)
  ) +
  labs(
    x="Code coverage", 
    y="Number of eval call sites", 
    size="Lines of R code",
    linetype=""
  ) +
  guides(
    size=guide_legend(order=1),
    color=FALSE,
    linetype=FALSE
  )

@@

```{r compute kaggle programs}
kaggle_programs <-
  kaggle %>%
  filter(exitval==0) %>%
  transmute(
    id, 
    code, 
    runtime=as.numeric(runtime, units="seconds")
  )
```

```{r kaggle summary}
kaggle_code     <- kaggle_programs$code
kaggle_runtime  <- kaggle_programs$runtime

overview_table(
  r("kaggle",                 kaggle),
  r("duplicated kaggle",      nrow(kaggle)-nrow(filter(kaggle, runnable))),
  r("runnable kaggle",        filter(kaggle, runnable)),
  r("finished kaggle",        kaggle_programs),
  r("failed kaggle",          nrow(filter(kaggle, runnable))-nrow(kaggle_programs)),
  rs("finished kaggle code",   kaggle_code),
  rs("kaggle runtime",         as.duration(kaggle_runtime)),
)
```

@@

```{r compute programs}
programs <- bind_rows(
  corpus_programs %>% mutate(corpus="packages") %>% select(code, runtime, corpus),
  kaggle_programs %>% mutate(corpus="kaggle") %>% select(code, runtime, corpus)
)
```

>>>>

```{r load corpus}
evals_dynamic <- read_fst(PACKAGE_EVALS_DYNAMIC_FILE) %>% as_tibble() %>% rename(evals_dyn=n)

corpus <- 
  read_fst(CORPUS_DETAILS_FILE) %>%
  as_tibble() %>%
  left_join(evals_dynamic, by="package") %>%
  mutate_at(
    vars(
      starts_with("runnable_code"),
      starts_with("runnable_files"),
      starts_with("package_"),
      starts_with("funs_"),
      "evals",
      "evals_dyn"
    ), 
    replace_na, 
    0
  )
```

@@

```{r filter out inconsistent programs}
corpus_programs <-
  read_fst(CRAN_PROGRAMS_FILE) %>%
  as_tibble()
```

```{r summary}
runnable_code <- 
  corpus_programs %>%
  mutate(files=1) %>%
  pivot_wider(names_from=type, values_from=c(files, code)) %>%
  select(-file, -runtime, -error, -success, -corpus) %>%
  group_by(package) %>%
  summarise_all(sum, na.rm=TRUE) %>%
  group_by(package) %>%
  mutate(
    files=files_examples + files_vignettes,
    code=code_examples + code_vignettes
  ) %>%
  ungroup()

@@

  rs("package programs",         runnable_code$files),
  rs("package programs code",    runnable_code$code),

  rs("examples programs",  runnable_code$files_examples),
  rs("vignettes programs", runnable_code$files_vignettes),

  rs("examples code",  corpus_program_examples$code),
  rs("vignettes code", corpus_programs_vignettes$code),

@@

corpus %>%
  mutate(
    outlier=evals_dyn > PACKAGE_EVALS_DYN_OUTLIER,# | package_r_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(outlier, package, NA),
    code_size=cut(package_r_code, size_bins, labels=size_bins_lasb, include.lowest=TRUE)
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
#      y=funs_with_eval,
      y=evals_dyn,
      label=label,
      size=code_size
    )
  ) +
  geom_point(alpha=.4) +
  geom_text(size=3, check_overlap = T, vjust=-1.5, na.rm = TRUE) + 
  geom_hline(aes(yintercept=funs_with_eval_mean), linetype="dashed", color="red", size=0.2) +
  geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="red", size=0.2) +
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(trans=scales::pseudo_log_trans(), breaks=c(0, 1, 2, 5, 10, 20, 50, 100, 200, 500)) +
  scale_size_discrete(range = c(1, 6)) +
  theme(
    legend.position=c(0.2, 0.85),
    legend.box="horizontal",
    legend.box.background = element_rect(fill="white", size=0.1)
  ) +
  labs(
    x="Code coverage", 
    y="Number of eval call sites", 
    size="Lines of R code",
    linetype=""
  ) +
  guides(
    size=guide_legend(order=1),
    color=FALSE,
    linetype=FALSE
  )

@@

```{r kaggle dependencies}
kaggle_dependencies <- tibble(package=read_lines(KAGGLE_DEPENDENCIES_FILE))
```

```{r compute kaggle programs}
kaggle <- read_fst(KAGGLE_KERNELS_FILE) %>% as_tibble()
kaggle_programs <-
  read_fst(KAGGLE_PROGRAMS_FILE) %>% as_tibble() %>% filter(success)
```

```{r kaggle summary}
kaggle_code     <- kaggle_programs$code

overview_table(
  r("kaggle",                 kaggle),
  r("duplicated kaggle",      nrow(kaggle)-nrow(filter(kaggle, runnable))),
  r("runnable kaggle",        filter(kaggle, runnable)),
  r("finished kaggle",        kaggle_programs),
  r("failed kaggle",          nrow(filter(kaggle, runnable))-nrow(kaggle_programs)),
  rs("finished kaggle code",  kaggle_code)
)
```

```{r}
setdiff(kaggle_dependencies$package, corpus$package)
```

@@

```{r compute programs}
programs <- bind_rows(
  corpus_programs %>% select(code, runtime, corpus),
  kaggle_programs %>% select(code, runtime, corpus)
)
```
